\mytitle

### 2012-03-15 papers/SAS2012/implicitization.tex
mise à jour de l'exemple

This new subgraph is the union of the paths of the previous one with a set
$P'$ of new paths that become feasible regarding the current abstract values.
The paths in $P'$ are computed one after another, until no more path
can make the invariant grow. This is line~\ref{alg=computeNewPaths} in
Algorithm~\ref{algo:combined}, which corresponds to
Algorithm~\ref{algo:computepaths}. We also use SMT solving to discover
these new paths, but we subtly change the SMT formula given to the
SMT solver: we now try to find a path that is not yet in $P$, but is feasible
and makes the invariant candidate of its destination grow.
% À changer si on décide de ne plus distinguer $X^s$ et $X^d$
We thus check the satisfiability of the formula
$f'(p_i)$, where:
\vspace{-.5em}
\[
f'(p_i) = f(p_i)\wedge \neg P
\vspace{-1em}
\]
$X_j$ is updated using an abstract union when the point $p_j$ is the target of a new path. This way, further SMT queries do not
compute other paths with the same source and destination if it is not needed
(because these new paths would not make $X_j$ grow, hence would not be returned
by the SMT solver).

@@

\subsection{Termination}
Termination of this algorithm is guaranteed, because:
\begin{inparaenum}
\item 
the subset of paths $P$ strictly increases at each loop iteration,
and is bounded by the finite set of
paths in the entire graph. 
\item
	When computing new paths, we cunjunct our formula with $\neg P$, meaning
	that we obtain each possible path only once. The number of path is finite,
	so this computation always terminates.
\item
	The Path-focusing iterations terminate because of the properties of
	widening.
\end{inparaenum}

@@

We choose $P_R$ as the set of loop headers of the function, plus the initial
state. In this case, we have three elements in $P_R$.
We also unroll the loop once, in order to distinguish the paths that do not go
through the loop from the others:
\lstinline|if (wait()) {while (wait()) {}}|

The main loop in the expanded multigraph has then 8 distinct paths, half of
them being self loops, and the other half going to the header of the nested loop.

Guided static analysis from \cite{DBLP:conf/sas/GopanR07} yields, at line 5,
$\lstinline|x_old| \in (-\infty,+\infty)$.
Path-focusing \cite{Monniaux_Gonnord_SAS11} performs slightly
better, and finds $\lstinline|x_old| \in (-\infty,10000]$.
Now, let us see how our technique performs on this example.

@@

\begin{compactenum}
	\item The first subset of paths is depicted on Figure
		\ref{fig:example-graph} Step 1. We apply Path-Focusing on this graph,
		and discover for $p_5$ the inductive invariant $\lstinline|x_old| = 0$.
	\item
		We then compute the set $P'$ of paths that have to be added into the
		next subgraph. 
		The image of $\lstinline|x_old| = 0$ by the path that goes from $p_5$ to
		itself, and that goes through the \emph{else} branch of each 
		\emph{if-then-else}, is $-10 \leq \lstinline|x_old| \leq 10$. This path
		is then added to our subgraph. Moreover, there is no other path whose
		image is not in $-10 \leq \lstinline|x_old| \leq 10$.
		Additionally, the path that goes from $p_5$ to
		$p_{10}$, that goes through the \emph{else} branch of each 
		\emph{if-then-else} at lines 7, 8 and 9, is also 
		$-10 \leq \lstinline|x_old| \leq 10$, and thus is also added into the
		new subgraph. Again, there is no new path that goes to $p_{10}$ with an
		image outside $-10 \leq \lstinline|x_old| \leq 10$.
	\item
		We apply Path Focusing on this new subgraph (\ref{fig:example-graph}
		Step 2). When we have reached an inductive invariant, we narrow, and the
		result for both $p_5$ and $p_{10}$ is the polyhedron 
		$-10000 \leq \lstinline|x_old| \leq 10000$.
	\item Finally, we compute the new subgraph. The SMT-solver does not find any
		new path that makes the abstract values grow, and the algorithm
		terminates.
\end{compactenum}

Our technique gives us the expected invariant 
$\lstinline|x_old| \in~[-10000,\allowbreak 10000]$. 
Here, only 3 paths out of the 8 have been computed during the analysis. In
practice, depending on the order the SMT-solver returns the paths, other
feasible paths could have been added during the analysis.

\begin{figure}
\begin{minipage}[c]{.15\textwidth}
\begin{tikzpicture}[->,>=stealth',auto,node distance=2.5cm,
                    semithick,font=\footnotesize]

	\node[PRstate] (n00) {$p_4$};
	\node[PRstate] (n0) [below of=n00, yshift=0.8cm] {$p_5$};
	\node[PRstate] (n1) [below of=n0] {$p_{10}$};
	\node (label) [left of=n00, node distance=1cm] {\bf Step 1};

  \path [transition] 
		(n00) edge  node {$\lstinline|x_old| \gets 0$} (n0);
\end{tikzpicture}
\end{minipage}\hfill
\begin{minipage}[c]{.75\textwidth}
\begin{tikzpicture}[->,>=stealth',auto,node distance=2.1cm,
                    semithick,font=\footnotesize]

	\node[PRstate] (n00) {$p_4$};
	\node[PRstate] (n0) [below of=n00, yshift=0.8cm] {$p_5$};
	\node[PRstate] (n1) [below of=n0] {$p_{10}$};
	\node (label) [left of=n00, node distance=1cm] {\bf Step 2};

  \path [transition] 
		(n00) edge  node {$\lstinline|x_old| \gets 0$} (n0);
  \path [transition] 
		(n0) edge  [loop left] node [left, xshift=0.2cm] {
		$\begin{array}{r}
			-10000 \leq \lstinline|x| \leq 10000 \\
			\lstinline|x_old|-10 \leq \lstinline|x| \leq \lstinline|x_old|+10 /\\
			\lstinline|x_old| \gets \lstinline|x|
		\end{array}$
		} (n0);
  \path [transition] 
		(n0) edge node [left, yshift=-0.2cm, xshift=0] {
		$\begin{array}{r}
			-10000 \leq \lstinline|x| \leq 10000 \\
			\lstinline|x_old|-10 \leq \lstinline|x| \leq \lstinline|x_old|+10 /\\
			\lstinline|x_old| \gets \lstinline|x|
		\end{array}$
		} (n1);
\end{tikzpicture}
\end{minipage}
\caption{Ascending sequence of subgraphs}
\label{fig:example-graph}
\end{figure}

In this example, we see that our technique actually combines best of
\emph{Guided Static Analysis} and \emph{Path Focusing}. Still, we need to unroll
the loop once in order to get precise results. Section \ref{sec:disjunctive}
proposes a new technique that gives precise invariants without needing such
graph transformations.

>>>>

This new subgraph contains all the paths from the previous one, and also new
paths that become feasible regarding the current abstract values.
The new paths in $P$ are computed one after another, until no more path
can make the invariant grow. This is line~\ref{alg=computeNewPaths} in
Algorithm~\ref{algo:combined}, which corresponds to
Algorithm~\ref{algo:computepaths}. We also use SMT solving to discover
these new paths, but we subtly change the SMT formula given to the
SMT solver: we now try to find a path that is not yet in $P$, but is feasible
and makes the invariant candidate of its destination grow.
% À changer si on décide de ne plus distinguer $X^s$ et $X^d$
We thus check the satisfiability of the formula
$f'(p_i)$, where:
\vspace{-.5em}
\[
f'(p_i) = f(p_i)\wedge \neg P
\vspace{-1em}
\]
$X_j$ is updated using an abstract union when the point $p_j$ is the target of a new path. This way, further SMT queries do not
compute other paths with the same source and destination if it is not needed
(because these new paths would not make $X_j$ grow, hence would not be returned
by the SMT solver).

@@

\subsection{Termination}
Termination of this algorithm is guaranteed, because:
\begin{inparaenum}
\item 
the subset of paths $P$ strictly increases at each loop iteration,
and is bounded by the finite set of
paths in the entire graph. 
\item
	when computing new paths, we cunjunct our formula with $\neg P$, meaning
	that we obtain each possible path only once. The number of path is finite,
	so this computation always terminates.
\item
	the Path-focusing iterations terminate because of the properties of
	widening.
\end{inparaenum}

@@

We choose $P_R$ as the set of loop headers of the function, plus the initial
state. In this case, we have three elements in $P_R$.

The main loop in the expanded multigraph has then 4 distinct paths going to the header of the nested loop.

Guided static analysis from \cite{DBLP:conf/sas/GopanR07} yields, at line 3,
$\lstinline|x_old| \in (-\infty,+\infty)$.
Path-focusing \cite{Monniaux_Gonnord_SAS11} also
finds $\lstinline|x_old| \in (-\infty,+\infty)$.
Now, let us see how our technique performs on this example.

@@

\begin{compactenum}
	\item The starting subgraph is depicted on Figure
		\ref{fig:example-graph} Step 1. At the beginning, this graph has no
		transitions.
	\item
		We compute the new feasible paths that have to be added into the
		subgraph. We first find the path from $p_1$ to $p_3$ and obtain at $p_3$
		$\lstinline|x_old| = 0$. We then find the path

		The image of $\lstinline|x_old| = 0$ by the path that goes from $p_3$ to
		$p_8$, and that goes through the \emph{else} branch of each 
		\emph{if-then-else}, is $-10 \leq \lstinline|x_old| \leq 10$. This path
		is then added to our subgraph.

		Moreover, there is no other path starting at $p_3$ whose
		image is not in $-10 \leq \lstinline|x_old| \leq 10$.

		Finally, since the abstract value associated to $p_8$ is  
		$-10 \leq \lstinline|x_old| \leq 10$, the path from $p_8$ to $p_1$ is
		feasible and is added into $P$. The final subgraph is depicted on Figure
		\ref{fig:example-graph} Step 2.

	\item
		We then compute the ascending iterations by path-focusing. At the end of
		these iterations, we obtain 
$-\infty \leq \lstinline|x_old| \leq +\infty$ for both $p_3$ and $p_8$.
	\item
		We now can apply narrowing iterations, and recover the precision lost by
		widening: we obtain 
		$-10000 \leq \lstinline|x_old| \leq 10000$ at points $p_3$ and $p_8$.
	\item Finally, we compute the next subgraph. The SMT-solver does not find any
		new path that makes the abstract values grow, and the algorithm
		terminates.

\end{compactenum}

Our technique gives us the expected invariant 
$\lstinline|x_old| \in~[-10000,\allowbreak 10000]$. 
Here, only 3 paths out of the 6 have been computed during the analysis. In
practice, depending on the order the SMT-solver returns the paths, other
feasible paths could have been added during the analysis.

\begin{figure}
\begin{minipage}[c]{.45\textwidth}
\begin{tikzpicture}[->,>=stealth',auto,node distance=2.1cm,
                    semithick,font=\footnotesize]

	\node[PRstate] (n00) {$p_1$};
	\node[PRstate] (n0) [below of=n00, yshift=0.8cm] {$p_3$};
	\node[PRstate] (n1) [below of=n0] {$p_{8}$};
	\node (label) [left of=n00, node distance=1cm] {\bf Step 1};

\end{tikzpicture}
\end{minipage}\hfill
\begin{minipage}[c]{.75\textwidth}
\begin{tikzpicture}[->,>=stealth',auto,node distance=2.1cm,
                    semithick,font=\footnotesize]

	\node[PRstate] (n00) {$p_1$};
	\node[PRstate] (n0) [below of=n00, yshift=0.8cm] {$p_3$};
	\node[PRstate] (n1) [below of=n0] {$p_{8}$};
	\node (label) [left of=n00, node distance=1cm] {\bf Step 2};

  \path [transition] 
		(n00) edge  node {$\lstinline|x_old| \gets 0$} (n0);
  \path [transition] 
		(n0) edge [bend left] node [left, yshift=-0.2cm, xshift=0] {
		$\begin{array}{r}
			-10000 \leq \lstinline|x| \leq 10000 \\
			\lstinline|x_old|-10 \leq \lstinline|x| \leq \lstinline|x_old|+10 /\\
			\lstinline|x_old| \gets \lstinline|x|
		\end{array}$
		} (n1);
  \path [transition] 
		(n1) edge [bend left]  node {} (n0);
\end{tikzpicture}
\end{minipage}
\caption{Ascending sequence of subgraphs}
\label{fig:example-graph}
\end{figure}

In this example, we see that our technique actually combines best of
\emph{Guided Static Analysis} and \emph{Path Focusing}. 
% JH: plus besoin de ca
%Still, we need to unroll
%the loop once in order to get precise results. Section \ref{sec:disjunctive}
%proposes a new technique that gives precise invariants without needing such
%graph transformations.
